---
title: "Results Simulation Study"
author: "Florian Stijven"
date: "`r Sys.Date()`"
output: bookdown::html_document2
bibliography: references.bib
---

```{r setup, include=FALSE}
# Set global markdown options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
# Load libraries that will be used further on.
library(ggplot2)
library(dplyr)
library(TCT)
```

# Introduction

In this Rmarkdown document, we analyze the results of the simulation study for
assessing the finite sample properties of the meta TCT methods. The primary goal
of this simulation study is to asses to what degree the theoretical asymptotic
results translate to finite samples. The secondary goal is to assess the
correctness of the implementation in the `TCT` R-package.

All simulated data sets were generated from multivariate normal distributions 
matching the data generating model (DGM) used by @raket_progression_2022 [Section 5.1].
This DGM represents a realistic 36-months clinical trial in
prodromal Alzheimer's disease. The control group is based on the analysis of a 
selection[^1] of 556 patients from the Alzheimer's disease neuroimaging initiative 
[@veitch2019understanding].

[^1]: The following inclusing criteria were used by @raket_progression_2022: "at
the baseline visit, patients must be diagnosed as having mild cognitive
impairment, score less than or equal to 28 on the mini mental state examination
(MMSE, range 30-0, higher scores indicate less impairment) and be amyloid
positive according to a brain positron emission tomography scan or analysis of
cerebrospinal fluid."

We first present the DGM in more details together with a visualization of the
relevant mean trajectories. Second, the analysis methods we consider in this
simulation study are summarized. Finally, the simulation results are presented.

# Data Generating Model

For the 556 patients selected from the Alzheimer's disease neuroimaging initiative,
the ADAS-cog scores were[^2] available at baseline visits and 6, 12, 18, 24, and 36
months after baseline. The estimated means at the corresponding time points are
$$(19.6, 20.5, 20.9, 22.7, 23.8, 27.4)'$$
and the corresponding estimated covariance matrix is 
$$\begin{pmatrix}
45.1 & 40.0 & 45.1 & 54.9 & 53.6 & 60.8 \\
40.0 & 57.8 & 54.4 & 66.3 & 64.1 & 74.7 \\
45.1 & 54.4 & 72.0 & 80.0 & 77.6 & 93.1 \\
54.9 & 66.3 & 80.0 & 109.8 & 99.3 & 121.7 \\
53.6 & 64.1 & 77.6 & 99.3 & 111.4 & 127.8 \\
60.8 & 74.7 & 93.1 & 121.7 & 127.8 & 191.4
\end{pmatrix}.$$
Patient-level data in the control group are generated from a multivariate normal
distribution with the above mean vector and covariance matrix, as in 
@raket_progression_2022 [Section 5.1]. Whereas @raket_progression_2022 considered
multiple types of treatment effects, we only consider treatment effects that 
correspond to proportional slowing. 

To simulate proportional slowing, we first consider the reference trajectory,
that is, the mean ADAS-cog score in the control group ($Z = 0$) as a function of
time since baseline ($t$), $$f_{0}(t; \boldsymbol{\alpha}) = E(Y_{t} | Z = 0)$$
where $Y_t$ is the ADAScog score $t$ months after baseline. This trajectory
should be a continuous function defined for every time point in the relevant
range, $[0, 36]$. Because we only have 6 mean estimates, at 6 distinct time
points, we interpolate between these 6 points with natural cubic interpolation.
This *interpolated* reference trajectory is given in Figure
\@ref(fig:data-generating-model-visualization).

[^2]: 13-item cognitive subscale of the Alzheimer’s disease assessment scale, 
lower scores indicate less impairment [@rosen1984new]

For simulating data from the treated group ($Z = 1$), we consider trajectories of the
following form, $$E(Y_t | Z = 1) = f_0(\gamma \cdot t; \boldsymbol{\alpha}) $$
where $\gamma$ is the proportional slowing factor. These data are simulated from
a multivariate normal distribution with the means determined by the above
trajectory function and with the same covariance matrix as in the control group.

We consider a set of DGMs where the following elements are varied to represent
a range of realistic scenarios:

* **Progression Rate**. We consider a *normal* and *fast* progression rate. The 
normal progression rate corresponds to the mean vector presented above. For the 
fast progression rate, we change the mean vector in the control group to 
$$(18.0, 19.7, 20.9, 22.7, 24.7, 29.2)'.$$ Alternatively, the fast progression 
scenario could also be interpreted as corresponding to trials where patients have
been followed up longer.
* **Treatment Effect**. We consider 3 different (proportional slowing) treatment 
effects, that is, $\gamma \in \{1, 0.75, 0.50 \}$.
* **Sample Size**. We consider 4 different total sample sizes, $n \in \{50, 200, 500, 1000\}$.
Note that $n$ is the total sample size, and we assume $1:1$ randomization in all settings.
* **Duration of follow up**. We consider settings with 24 and 36 months of follow up, 
corresponding to 5 and 6 measurements of the ADAScog score, respectively.

```{r data-generating-model-visualization, fig.cap = "Plot of the trajectories used in the DGMs. A slowing factor equal to 1 corresponds to no treatment effect."}
ref_means_list = list(
  "Normal Progression" = c(19.6, 20.5, 20.9, 22.7, 23.8, 27.4),
  "Fast Progression" = c(18, 19.7, 20.9, 22.7, 24.7, 29.2)
)
time_points = c(0, 6, 12, 18, 24, 36)
time_grid = seq(from = 0,
                to = 36,
                length.out = 3000)
dgm_settings = tidyr::expand_grid(
  progression = c("Normal Progression", "Fast Progression"),
  gamma_slowing = c(1, 0.75, 0.5),
  time_points = list(time_points)
)
trajectory_observed_tbl = dgm_settings %>%
  rowwise(everything()) %>%
  summarise(ref_means = ref_means_list[progression],
            trajectory_points = list(
              spline(
                x = time_points,
                y = ref_means,
                xout = gamma_slowing * unlist(time_points)
              )$y
            )) %>%
  ungroup() %>%
  mutate(treatment = ifelse(gamma_slowing == 1, "Control Treatment", "Active Treatment"))
trajectory_observed_tbl = trajectory_observed_tbl %>%
  rowwise(everything()) %>%
  reframe(tibble(
    trajectory_points = unlist(trajectory_points),
    time_points = unlist(time_points)
  )) %>%
  ungroup() %>%
  mutate(gamma_slowing = as.factor(gamma_slowing))

trajectory_interpolated_tbl = dgm_settings %>%
  rowwise(everything()) %>%
  summarise(
    ref_means = ref_means_list[progression],
    trajectory_interpolated = list(
      spline(
        x = time_points,
        y = ref_means,
        xout = gamma_slowing * time_grid
      )$y
    ),
    time_grid = list(time_grid)
  ) %>%
  ungroup() %>%
  mutate(treatment = ifelse(gamma_slowing == 1, "Control Treatment", "Active Treatment"))
trajectory_interpolated_tbl = trajectory_interpolated_tbl %>%
  rowwise(everything()) %>%
  reframe(tibble(
    trajectory_interpolated = unlist(trajectory_interpolated),
    time_grid = unlist(time_grid)
  )) %>%
  ungroup() %>%
  mutate(gamma_slowing = as.factor(gamma_slowing))

trajectory_observed_tbl %>%
  ggplot(aes(x = time_points, y = trajectory_points)) +
  geom_point(alpha = 0.25) +
  geom_line(
    data = trajectory_interpolated_tbl,
    mapping = aes(
      x = time_grid,
      y = trajectory_interpolated,
      group = interaction(gamma_slowing, treatment, progression),
      color = gamma_slowing
    )
  ) +
  scale_x_continuous(breaks = time_points) +
  scale_color_discrete(type = "seq", name = latex2exp::TeX("Slowing Factor, \\gamma")) +
  xlab("Months After Randomization") +
  ylab("Mean Trajectory") +
  facet_grid( ~ progression)
```

# Analysis Methods

In this section, two analysis methods are briefly outline. We first present the 
mixed model for repeated measures (MMRM) that is fitted to the simulated data sets.
Next, we discuss the meta TCT methods which use the fitted MMRM. 

## Mixed Model for Repeated Measures

For each simulated data set, a MMRM is fitted. This is a linear mixed model
where time is treated as a categorical covariate. The **systematic part** consists
of the interaction between treatment and time, except for $t = 0$ where we
assume that the mean outcome is equal in both treatment groups. Let $j = 0, 1, ..., K$
denote the measurement occasions, and $t_j$ the corresponding months after baseline.
The mean outcome is then modeled as
$$E(Y_{t_j} | Z = z) = \beta_{z, j} \; \forall \; j \in \{ 1, ..., K \}$$
and $E(Y_{0} | Z = 0) = E(Y_{0} | Z = 1) = \beta_0$.
This essentially
means that we have a parameter for each measurement occasion-treatment
combination, except for baseline. The **covariance matrix** is assumed to be
unstructured, but common for both treatment groups.
For all simulation scenarios, this is a correctly specified model. 

These models are fitted using restricted maximum likelihood (REML) using the
`mmrm()` function from the `mmrm` R-package [@mmrmpackage]. The parameter
estimates and corresponding variance-covariance matrix are obtained by the
`coef()` and `vcov()` methods. The hypothesis test for 
$$H_0: \beta_{0, j} = \beta_{1, j} \; \forall \; j \in \{ 1, ..., K \} $$
is based on the F-test with the Kenward-Roger approximate degrees of freedom.

## Meta TCT

TO DO

# Results

In this section, the results of the simulation study are presented. For each 
setting, we considered $N$ replications. For estimating coverage of 95\% CIs and
empirical type 1 errors, this leads to a standard error of $X$. 

First, we present the results regarding estimation in terms of 



Second, we look at operating characteristics of the inference procedures. These 
also compared with inference based on the MMRM.

```{r}
# Read data set with results of the simulation study.
results_tbl = readRDS(file = "results_simulation_lean.rds") %>%
  # We also extract the estimated acceleration factor into a seperate column.
  mutate(estimate = purrr::map_dbl(
    .x = TCT_meta_common_fit,
    .f = coef
  ))
```


## Estimation

```{r}
# Compute all summary measures regarding the performance of the estimators.
results_tbl_estimation = results_tbl %>%
  # Group by all simulation scenarios.
  group_by(progression, gamma_slowing, n, K, drop_first_occasions, constraints, inference) %>%
  summarise(mean_estimate = mean(estimate),
            empirical_sd = sd(estimate),
            mean_se = mean(se_TCT_common),
            mse = mean((estimate - gamma_slowing)**2)) %>%
  mutate("Follow Up" = ifelse(K == 6, "36 Months", "24 Months"))
```

In Figure X, we present the results regarding estimation across all scenarios.

```{r expected-value-estimator, fig.cap = "Graph of estimated expected value of the estimator for the common accleration factor across all simulation settings."}
results_tbl_estimation %>%
  filter(constraints == TRUE, inference == "least-squares") %>%
  ggplot(aes(x = n, y = mean_estimate, color = `Follow Up`, linetype = progression)) +
  geom_point() + 
  geom_line() + 
  geom_hline(
    # Add horizontal lines to indicate the true values.
    data = tidyr::expand_grid(gamma_slowing = c(0.5, 0.75, 1), 
                              drop_first_occasions = 0:2),
    mapping = aes(
      yintercept = gamma_slowing
    ),
    alpha = 0.5) +
  facet_grid(gamma_slowing~drop_first_occasions, scales = "free")
```

```{r mse-estimator, fig.cap = "Graph of the means squared error of the estimator for the common accleration factor across all simulation settings."}
results_tbl_estimation %>%
  filter(constraints == TRUE, inference == "least-squares") %>%
  ggplot(aes(x = n, y = mse, color = `Follow Up`, linetype = progression)) +
  geom_point() + 
  geom_line() +
  scale_y_continuous(trans = "log10") +
  scale_x_continuous(trans = "log10") +
  facet_grid(gamma_slowing~drop_first_occasions, scales = "free")
```

```{r se-estimator, fig.cap = "Graph of the empirical standard deviations and the mean estimated standard error of the estimator for the common accleration factor across all simulation settings."}
results_tbl_estimation %>%
  filter(constraints == TRUE, inference == "least-squares") %>%
  ggplot(aes(
    x = n,
    y = empirical_sd,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  geom_point(
    data = results_tbl_estimation %>%
      filter(constraints == TRUE, inference == "least-squares"),
    mapping = aes(x = n, y = mean_se),
    shape = 2
  ) +
  scale_y_continuous(trans = "log10") +
  scale_x_continuous(trans = "log10") +
  facet_grid(gamma_slowing ~ drop_first_occasions, scales = "free")
```





## Inference


```{r}
# Compute all summary measures regarding the performance of the estimators.
results_tbl_inference = results_tbl %>%
  # Group by all simulation scenarios.
  group_by(progression, gamma_slowing, n, K, drop_first_occasions, constraints, inference) %>%
  summarise(mean_estimate = mean(estimate),
            empirical_sd = sd(estimate),
            mean_se = mean(se_TCT_common),
            mse = mean((estimate - gamma_slowing)**2)) %>%
  mutate("Follow Up" = ifelse(K == 6, "36 Months", "24 Months"))
```

```{r}
results_tbl_inference = results_tbl %>%
  # Extract lower and upper 95% confidence limits into two new columns.
  mutate(lower_conf_int = purrr::map_dbl(.x = conf_int_TCT_common, .f = 1),
         upper_conf_int = purrr::map_dbl(.x = conf_int_TCT_common, .f = 2)) %>%
  mutate(
    # Determine whether the CI contain the true value.
    covered = (gamma_slowing <= upper_conf_int) & (gamma_slowing >= lower_conf_int),
    # Determine whether the null has been rejected
    rejection = p_value_TCT_common <= 0.05,
    rejection_mmrm = p_value_mmrm <= 0.05
    ) %>%
  group_by(progression, gamma_slowing, n, K, drop_first_occasions, constraints, inference) %>%
  summarise(
    coverage = mean(covered),
    rejection_rate = mean(rejection),
    rejection_rate_mmrm = mean(rejection_mmrm)) %>%
  mutate("Follow Up" = ifelse(K == 6, "36 Months", "24 Months"))
```

```{r}
results_tbl_inference %>%
  filter(constraints == TRUE, inference == "least-squares") %>%
  ggplot(aes(
    x = n,
    y = rejection_rate,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  facet_grid(gamma_slowing ~ drop_first_occasions, scales = "free")
```

```{r}
results_tbl_inference %>%
  filter(constraints == TRUE, inference == "least-squares") %>%
  ggplot(aes(
    x = n,
    y = rejection_rate_mmrm,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  facet_grid(gamma_slowing ~ drop_first_occasions, scales = "free")
```

```{r}
results_tbl_inference %>%
  filter(constraints == TRUE, inference == "least-squares") %>%
  ggplot(aes(
    x = n,
    y = coverage,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  facet_grid(gamma_slowing ~ drop_first_occasions, scales = "free")
```





