---
title: "Results Simulation Study"
author: "Florian Stijven"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    toc: true
bibliography: references.bib
---

```{r setup, include=FALSE}
# Set global markdown options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
# Load libraries that will be used further on.
library(ggplot2)
library(dplyr)
library(TCT)
```

# Introduction

In this Rmarkdown document, we present the results of the simulation study for
assessing the finite sample properties of the meta TCT methods. The primary goal
of this simulation study is to asses to what degree the theoretical asymptotic
results translate to finite samples. The secondary goal is to identify finite
sample settings where the meta TCT methods are not trustworthy. The tertiary
goal is to assess the correctness of the implementation in the `TCT` R-package.

All simulated data sets were generated from multivariate normal distributions 
matching the data generating model (DGM) used by @raket_progression_2022 [Section 5.1].
This DGM represents a realistic 36-months clinical trial in
prodromal Alzheimer's disease. The control group is based on the analysis of a 
selection[^1] of 556 patients from the Alzheimer's disease neuroimaging initiative 
[@veitch2019understanding].

[^1]: The following inclusion criteria were used by @raket_progression_2022: "at
the baseline visit, patients must be diagnosed as having mild cognitive
impairment, score less than or equal to 28 on the mini mental state examination
(MMSE, range 30-0, higher scores indicate less impairment) and be amyloid
positive according to a brain positron emission tomography scan or analysis of
cerebrospinal fluid."

We first present the DGM in more details together with a visualization of the
relevant mean trajectories. Second, the analysis methods we consider in this
simulation study are summarized. Finally, the simulation results are presented.

# Data Generating Model

For the 556 patients selected from the Alzheimer's disease neuroimaging initiative,
the ADAS-cog scores were[^2] available at baseline visits and 6, 12, 18, 24, and 36
months after baseline. The estimated means at the corresponding time points are
$$(19.6, 20.5, 20.9, 22.7, 23.8, 27.4)'$$
and the corresponding estimated covariance matrix is 
$$\begin{pmatrix}
45.1 & 40.0 & 45.1 & 54.9 & 53.6 & 60.8 \\
40.0 & 57.8 & 54.4 & 66.3 & 64.1 & 74.7 \\
45.1 & 54.4 & 72.0 & 80.0 & 77.6 & 93.1 \\
54.9 & 66.3 & 80.0 & 109.8 & 99.3 & 121.7 \\
53.6 & 64.1 & 77.6 & 99.3 & 111.4 & 127.8 \\
60.8 & 74.7 & 93.1 & 121.7 & 127.8 & 191.4
\end{pmatrix}.$$
Patient-level data in the control group are generated from a multivariate normal
distribution with the above mean vector and covariance matrix, as in 
@raket_progression_2022 [Section 5.1]. Whereas @raket_progression_2022 considered
multiple types of treatment effects, we only consider treatment effects that 
correspond to proportional slowing. 

To simulate proportional slowing, we first consider the reference trajectory,
that is, the mean ADAS-cog score in the control group ($Z = 0$) as a function of
time since baseline ($t$), $$E(Y_{t} | Z = 0) = f_{0}(t; \boldsymbol{\alpha})$$
where $Y_t$ is the ADAScog score $t$ months after baseline. This trajectory
should be a continuous function defined for every time point in the relevant
range, $[0, 36]$. Because we only have 6 mean estimates, at 6 distinct time
points, we interpolate between these 6 points with natural cubic spline
interpolation. This *interpolated* reference trajectory is plotted in Figure
\@ref(fig:data-generating-model-visualization).

[^2]: 13-item cognitive subscale of the Alzheimer’s disease assessment scale, 
lower scores indicate less impairment [@rosen1984new]

For simulating data from the treated group ($Z = 1$), we consider trajectories of the
following form, $$E(Y_t | Z = 1) = f_0(\gamma \cdot t; \boldsymbol{\alpha}) $$
where $\gamma$ is the proportional slowing factor. These data are simulated from
a multivariate normal distribution with the means determined by the above
trajectory function and with the same covariance matrix as in the control group.

We consider a set of DGMs where the following elements are varied to represent
a range of realistic scenarios:

* **Progression Rate**. We consider a *normal* and *fast* progression rate. The 
normal progression rate corresponds to the mean vector presented above. For the 
fast progression rate, we change the mean vector in the control group to 
$$(18.0, 19.7, 20.9, 22.7, 24.7, 29.2)'.$$ Alternatively, the fast progression 
scenario could also be interpreted as corresponding to trials where patients have
been followed up longer.
* **Treatment Effect**. We consider 3 different (proportional slowing) treatment 
effects, that is, $\gamma \in \{1, 0.75, 0.50 \}$.
* **Sample Size**. We consider 4 different total sample sizes, $n \in \{50, 200, 500, 1000\}$.
Note that $n$ is the *total* sample size, and we assume $1:1$ randomization in all settings.
* **Duration of follow up**. We consider settings with 24 and 36 months of follow up, 
corresponding to 5 and 6 measurements of the ADAScog score, respectively.

```{r data-generating-model-visualization, fig.cap = "Plot of the trajectories used in the DGMs. A slowing factor equal to 1 corresponds to no treatment effect. The dots correspond to measurement occasions while the lines between dots are based on interpolation."}
ref_means_list = list(
  "Normal Progression" = c(19.6, 20.5, 20.9, 22.7, 23.8, 27.4),
  "Fast Progression" = c(18, 19.7, 20.9, 22.7, 24.7, 29.2)
)
time_points = c(0, 6, 12, 18, 24, 36)
time_grid = seq(from = 0,
                to = 36,
                length.out = 3000)
dgm_settings = tidyr::expand_grid(
  progression = c("Normal Progression", "Fast Progression"),
  gamma_slowing = c(1, 0.75, 0.5),
  time_points = list(time_points)
)
trajectory_observed_tbl = dgm_settings %>%
  rowwise(everything()) %>%
  summarise(ref_means = ref_means_list[progression],
            trajectory_points = list(
              spline(
                x = time_points,
                y = ref_means,
                xout = gamma_slowing * unlist(time_points)
              )$y
            )) %>%
  ungroup() %>%
  mutate(treatment = ifelse(gamma_slowing == 1, "Control Treatment", "Active Treatment"))
trajectory_observed_tbl = trajectory_observed_tbl %>%
  rowwise(everything()) %>%
  reframe(tibble(
    trajectory_points = unlist(trajectory_points),
    time_points = unlist(time_points)
  )) %>%
  ungroup() %>%
  mutate(gamma_slowing = as.factor(gamma_slowing))

trajectory_interpolated_tbl = dgm_settings %>%
  rowwise(everything()) %>%
  summarise(
    ref_means = ref_means_list[progression],
    trajectory_interpolated = list(
      spline(
        x = time_points,
        y = ref_means,
        xout = gamma_slowing * time_grid
      )$y
    ),
    time_grid = list(time_grid)
  ) %>%
  ungroup() %>%
  mutate(treatment = ifelse(gamma_slowing == 1, "Control Treatment", "Active Treatment"))
trajectory_interpolated_tbl = trajectory_interpolated_tbl %>%
  rowwise(everything()) %>%
  reframe(tibble(
    trajectory_interpolated = unlist(trajectory_interpolated),
    time_grid = unlist(time_grid)
  )) %>%
  ungroup() %>%
  mutate(gamma_slowing = as.factor(gamma_slowing))

trajectory_observed_tbl %>%
  ggplot(aes(x = time_points, y = trajectory_points)) +
  geom_point(alpha = 0.25) +
  geom_line(
    data = trajectory_interpolated_tbl,
    mapping = aes(
      x = time_grid,
      y = trajectory_interpolated,
      group = interaction(gamma_slowing, treatment, progression),
      color = gamma_slowing
    )
  ) +
  scale_x_continuous(breaks = time_points) +
  scale_color_discrete(type = "seq", name = latex2exp::TeX("Slowing Factor, \\gamma")) +
  xlab("Months After Randomization") +
  ylab("Mean Trajectory") +
  facet_grid( ~ progression)
```

# Analysis Methods

In this section, two analysis methods are briefly outlined. We first present the 
mixed model for repeated measures (MMRM) that is fitted to the simulated data sets.
Next, we discuss the meta TCT methods which use the fitted MMRMs as input. 

## Mixed Model for Repeated Measures

For each simulated data set, a MMRM is fitted. This is a linear mixed model
where time is treated as a categorical covariate. The **systematic part** consists
of the interaction between treatment and time, except for $t = 0$ where we
assume that the mean outcome is equal in both treatment groups. Let $j = 0, 1, ..., K$
denote the measurement occasions, and $t_j$ the corresponding months after baseline.
The mean outcome is then modeled as
$$E(Y_{t_j} | Z = z) = \beta_{z, j} \; \forall \; j \in \{ 1, ..., K \} \text{ and } z \in \{0, 1\}$$
and $E(Y_{0} | Z = 0) = E(Y_{0} | Z = 1) = \beta_0$.
This essentially
means that we have a parameter for each measurement occasion-treatment
combination, except for baseline. The **covariance matrix** is assumed to be
unstructured, but common for both treatment groups.
For all simulation scenarios, this is a correctly specified model. 

These models are fitted using restricted maximum likelihood (REML) using the
`mmrm()` function from the `mmrm` R-package [@mmrmpackage]. The parameter
estimates and corresponding variance-covariance matrix are obtained by the
`coef()` and `vcov()` methods. These two components are the only inputs required
for the meta TCT methods. The hypothesis test for 
$$H_0: \beta_{0, j} = \beta_{1, j} \; \forall \; j \in \{ 1, ..., K \} $$
is based on the F-test with the Kenward-Roger approximate degrees of freedom.

## Meta TCT

TO DO

# Results

In this section, the results of the simulation study are presented.
For each setting, we consider $5000$ replications. For estimating coverage of
95\% CIs and empirical type 1 errors, this leads to a standard error of
$\frac{\sqrt{0.05 \cdot 0.95}}{\sqrt{5000}} = 0.003$ assuming nominal coverage and a nominal
type 1 error.

In this section, we only present the results of the meta TCT estimator based on
generalized least squares that uses all measurement occasions. In the Appendix, 
the results for other estimators are presented in an interactive table. 


The analyses and presentation of the simulation results are divided into two 
parts. First, we present the results regarding
estimation of the acceleration factor. This is the primary goal of the meta TCT
methods: Transforming the treatment effect on difficult to interpret scales to 
the time scale.
Second, we present the results regarding inference, that is, hypothesis tests and
confidence intervals. The operating
characteristics of the meta TCT inference procedures are also compared with
those of MMRM.

```{r}
# Read data set with results of the simulation study.
results_tbl = readRDS(file = "results_simulation_lean.rds")
```


## Estimation

To evaluate estimation, we look at 3 key performance measures,

1. **Bias of the estimator.** We estimate $E(\hat{\gamma})$ 
across different settings and compare the estimated expectation with the 
true value, $\gamma_0$.
2. **Mean squared error (MSE).** The MSE is defined as
$E\{(\hat{\gamma} - \gamma_0)^2 \}$. This quantity is estimated as the mean of the 
squared differences between $\hat{\gamma}$ and $\gamma_0$. This quantity quantifies the average
distance between the estimator and the estimand, which depends on the variance and
bias.
3. **Empirical standard deviation.** The empirical standard deviation of the estimator simply
is the standard deviation of the estimator. This measure is estimated as the sample
standard deviation of the estimates in each setting. This value is compared with the
median *estimated standard error*.  


```{r}
# Compute all summary measures regarding the performance of the estimators.
results_tbl_estimation = results_tbl %>%
  # Group by all simulation scenarios.
  group_by(progression, gamma_slowing, n, K, drop_first_occasions, constraints, inference) %>%
  summarise(mean_estimate = mean(estimate),
            empirical_sd = sd(estimate),
            mean_se = mean(se_TCT_common),
            median_se = median(se_TCT_common),
            mean_se_bs = mean(se_TCT_common_bs),
            median_se_bs = median(se_TCT_common_bs),
            mse = mean((estimate - gamma_slowing)**2)) %>%
  mutate("Follow Up" = ifelse(K == 6, "36 Months", "24 Months"))
```

In Figure \@ref(fig:expected-value-estimator), the mean of the estimated acceleration factors
is presented across a set of scenarios. GENERAL BIAS? BIAS in SMALL SAMPLES?

```{r expected-value-estimator, fig.cap = "Graphs of the means of the estimated acceleration factors across a set of simulation settings. The presented results are based on least-squares meta TCT. The rows correspond to the true acceleration factor."}
# Define dummy tibble that allows us to set the axis limits in each facet
# separately.
facet_lims_tbl = tibble(
  n = NA,
  gamma_slowing = rep(c(0.5, 0.75, 1), each = 2),
  mean_estimate = c(0.45, 0.55, 0.70, 0.80, 0.95, 1.05),
  `Follow Up` = NA, 
  progression = NA
)
results_tbl_estimation %>%
  filter(
    constraints == FALSE,
    inference == "least-squares",
    drop_first_occasions == 0) %>%
      ggplot(
        aes(
          x = n,
          y = mean_estimate,
          color = `Follow Up`,
          linetype = progression
        )
      ) +
      geom_point() +
      geom_line() +
      geom_hline(
        # Add horizontal lines to indicate the true values.
        data = tidyr::expand_grid(
          gamma_slowing = c(0.5, 0.75, 1),
          drop_first_occasions = 0
        ),
        mapping = aes(yintercept = gamma_slowing),
        alpha = 0.5
      ) +
    geom_blank(data=facet_lims_tbl, aes(gamma_slowing, mean_estimate)) +
      xlab("Sample size, n") +
      ylab(latex2exp::TeX("E(\\hat{\\gamma})")) +
      scale_linetype(name = "Progression Rate") +
      facet_grid(gamma_slowing ~ ., scales = "free"
      )
```

In Figure \@ref(fig:mse-estimator), the MSE of the estimator for the common
acceleration factor is presented across the same set of scenarios as before. As
expected, the MSE decreases as a function of the sample size. Moreover, a longer
follow up and faster progression lead to a smaller MSE.

```{r mse-estimator, fig.cap = "Graphs of the MSEs for the estimator of the common accleration factor across a set of simulation settings. The presented results are based on the least-squares meta TCT. The rows correspond to the true acceleration factor. Note that both axes are log10-transformed."}
results_tbl_estimation %>%
  filter(constraints == FALSE,
         inference == "least-squares",
         drop_first_occasions == 0) %>%
  ggplot(aes(
    x = n,
    y = mse,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  scale_y_continuous(trans = "log10", name = "MSE") +
  scale_x_continuous(trans = "log10", name = "Sample Size, n") +
  scale_linetype(name = "Progression Rate") +
  facet_grid(gamma_slowing ~ .)
```

In Figure \@ref(fig:se-estimator), the empirical standard deviations are plotted 
together with the median estimated standard errors. For small sample sizes, the
standard error estimators underestimate the empirical standard error. However, 
this underestimation largely disappears for larger sample sizes.

Although inference is not based directly on the estimated standard error, this
*underestimation* would lead to issues if the estimated acceleration factor were
used in meta-analyses. We have therefore also implemented a bootstrap-based
estimator for the standard error, as presented in the next section.


```{r se-estimator, fig.cap = "Graphs of the empirical standard deviations and the median estimated standard errors of the estimator for the common acceleration factor across a set of simulation settings. The dots, and connecting lines, represent the empirical standard deviations. The presented results are based on least-squares meta TCT. The rows correspond to the true acceleration factor while the columns correspond to the duration of follow-up. The triangles represent the median estimated standard errors."}
results_tbl_estimation %>%
  filter(constraints == FALSE,
         inference == "least-squares",
         drop_first_occasions == 0) %>%
  ggplot(aes(
    x = n,
    y = empirical_sd,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  geom_point(
    data = results_tbl_estimation %>%
      filter(constraints == FALSE,
         inference == "least-squares",
         drop_first_occasions == 0),
    mapping = aes(x = n, y = median_se),
    shape = 2
  ) +
  scale_y_continuous(trans = "log10", name = "Empirical SD and median estimated SE") +
  scale_x_continuous(trans = "log10", name = "Sample Size, n") +
  scale_linetype(name = "Progression Rate") +
  facet_grid(gamma_slowing ~ `Follow Up`)
```



## Inference

To evaluate inference, we look at 2 key performance measures,

1. **Type 1 error and power.** We compare the empirical type 1 error with the
nominal rate, $\alpha = 0.05$. We do this for the meta TCT methods as well as for
the MMRM. The former is based on the latter, so we also expect that discrepancies
between empirical and nominal type 1 error rates for the MMRM will be reflected in
the meta TCT methods.
2. **Coverage.** We asses the empirical coverage rate of the estimated $95\%$ 
CIs. 


```{r}
results_tbl_inference = results_tbl %>%
  mutate(
    # Determine whether the CI contain the true value.
    covered = (gamma_slowing <= conf_int_TCT_common_upper) &
      (gamma_slowing >= conf_int_TCT_common_lower),
    covered_bs = (gamma_slowing <= conf_int_TCT_common_upper_bs) &
      (gamma_slowing >= conf_int_TCT_common_lower_bs),
    # Determine the conclusion based on a bootstrap-based CI test.
    rejection_bs = !((1 <= conf_int_TCT_common_upper_bs) &
                       (1 >= conf_int_TCT_common_lower_bs)
    ),
    # Determine whether the null has been rejected
    rejection = p_value_TCT_common <= 0.05,
    rejection_mmrm = p_value_mmrm <= 0.05
  ) %>%
  group_by(progression,
           gamma_slowing,
           n,
           K,
           drop_first_occasions,
           constraints,
           inference) %>%
  summarise(
    coverage = mean(covered),
    coverage_bs = mean(covered_bs),
    rejection_rate_bs = mean(rejection_bs),
    rejection_rate = mean(rejection),
    rejection_rate_mmrm = mean(rejection_mmrm)
  ) %>%
  mutate("Follow Up" = ifelse(K == 6, "36 Months", "24 Months"))
```

In Figure \@ref(fig:error-rates-meta-tct), we graph the empirical type 1 error rate and
power. The corresponding empirical operating characteristics for the F-test are superimposed 
in gray. 
This reveals that the type 1 error rate is inflated for ... Consequently, the power 
for the corresponding settings is not well-calibrated.
For ... settings, the empirical type 1 error is close to nominal, hence, the corresponding
empirical powers are well-calibrated and can be interpreted as usual. 
In these well-calibrated scenarios, the power of the meta TCT test is larger than 
the power of the corresponding F-test in the MMRM.

The tests presented in Figure \@ref(fig:error-rates-meta-tct) are asymptotic.
So, poor performance in small samples is not surprising. To provide trustworthy
inference, we also assess the performance of a parametric bootstrap in the next
section.

```{r error-rates-meta-tct, fig.cap = "Graphs of the empirical type 1 error rate and power across a set of simulation settings. The presented results are based on least-squares meta TCT. The rows correspond to the true acceleration factor while the columns correspond to the duration of follow-up."}
results_tbl_inference %>%
  filter(constraints == FALSE,
         inference == "least-squares",
         drop_first_occasions == 0) %>%
  ggplot(
    aes(
      x = n,
      y = rejection_rate,
      color = `Follow Up`,
      shape = `Follow Up`,
      linetype = progression
    )
  ) +
  geom_point() +
  geom_line() +
  geom_point(
    data = results_tbl_inference %>%
      filter(constraints == FALSE,
         inference == "least-squares",
         drop_first_occasions == 0),
    mapping = aes(x = n,
                  y = rejection_rate_mmrm,
                  shape = `Follow Up`),
    alpha = 0.5,
    color = "gray"
  ) +
  geom_line(
    data = results_tbl_inference %>%
      filter(constraints == FALSE,
         inference == "least-squares",
         drop_first_occasions == 0),
    mapping = aes(x = n,
                  y = rejection_rate_mmrm,
                  linetype = progression),
    alpha = 0.75,
    color = "gray"
  ) +
  geom_hline(
    # Add horizontal lines to indicate the 0.05 nominal error rate in the null
    # settings.
    data = tidyr::expand_grid(
      rejection_rate = 0.05,
      drop_first_occasions = 0,
      gamma_slowing = 1
    ),
    mapping = aes(yintercept = rejection_rate),
    alpha = 0.5
  ) +
  xlab("Sample size, n") +
  ylab(latex2exp::TeX("Empirical Power or Type 1 Error Rate")) +
  scale_linetype(name = "Progression Rate") +
  facet_grid(gamma_slowing ~ `Follow Up`, scales = "free")
```

In Figure \@ref(fig:coverage-meta-tct), the empirical coverage rates are
presented for the same settings as before. This reveals that there is
undercoverage for ... For ... settings, coverage is close to nominal.

As mentioned before, the coverage of alternative confidence intervals based on 
a parametric bootstrap is assessed in the next section.

```{r coverage-meta-tct, fig.cap = "Graphs of the empirical coverage across a set of simulation settings. The presented results are based on least-squares meta TCT. The rows correspond to the true acceleration factor while the columns correspond to the duration of follow-up."}
results_tbl_inference %>%
  filter(constraints == FALSE,
         inference == "least-squares",
         drop_first_occasions == 0) %>%
  ggplot(aes(
    x = n,
    y = coverage,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
    geom_hline(
    # Add horizontal lines to indicate the 0.95 nominal coverage.
    data = tidyr::expand_grid(
      coverage = 0.95,
      drop_first_occasions = 0:1,
      gamma_slowing = c(0.5, 0.75, 1)
    ),
    mapping = aes(yintercept = coverage),
    alpha = 0.5
  ) +
  xlab("Sample size, n") +
  ylab(latex2exp::TeX("Empirical Coverage")) +
  scale_linetype(name = "Progression Rate") +
  facet_grid(gamma_slowing ~ `Follow Up`)
```

# Parametric Bootstrap

In small samples, the operating characteristics of least-squares meta TCT deviate
considerably from nominal. This is not surprising because the corresponding 
inferential procedures are asymptotic. We provide an alternative inferential
procedure based on a parametric bootstrap. The idea behind this parametric 
bootstrap is to resample the estimated parameters from the MMRM from the
estimated multivariate normal sampling distribution, that is,
$$\hat{\boldsymbol{\beta}}^b \, \sim \, N\left( \hat{\boldsymbol{\beta}}, \hat{\Sigma}  \right)$$
where

* $\hat{\boldsymbol{\beta}}$ is the estimated mean vector.
* $\hat{\Sigma}$ is the estimate variance-covariance matrix of the sampling 
distribution.
* $\hat{\boldsymbol{\beta}}^b$ is the $b$'th bootstrap replicate of the mean
vector.

The $b$'th bootstrap replicate of the common acceleration factor,
$\hat{\gamma}^b$, is obtained by applying least squares meta TCT to
$\hat{\boldsymbol{\beta}}$ (and $\hat{\Sigma}$). Subsequent inference is based
on the $1 - \alpha$ percentile confidence interval, that is,
$$(\hat{\gamma}^b_{\alpha / 2}, \hat{\gamma}^b_{1 - \alpha / 2})$$ where
$\hat{\gamma}^b_{p}$ is the $p$-th percentile of the bootstrap distribution.

To limit the computational burden, we only use $B = 500$ bootstrap replications 
throughout. We also only consider the bootstrap for the normal progression 
scenarios because the largest deviations from nominal were observed there.

In Figure \@ref(fig:se-estimator-bootstrap), the empirical standard deviations
are plotted together with the median estimated standard errors that are based on
the parametric bootstrap. The median estimated standard errors now closely 
match the empirical standard deviation. So, the parametric bootstrap provides 
a standard error estimator that is valid even in very small samples. 

```{r se-estimator-bootstrap, fig.cap="Graphs of the empirical standard deviations and the median estimated standard errors of the estimator for the common acceleration factor. The standard errors are estimated thrrough the parametric bootstrap as explained in the text. The dots, and connecting lines, represent the empirical standard deviations. The rows correspond to the true acceleration factor. The triangles represent the median estimated standard errors."}
results_tbl_estimation %>%
  filter(
    constraints == FALSE,
    inference == "least-squares",
    progression == "normal",
    drop_first_occasions == 0
  ) %>%
  ggplot(aes(
    x = n,
    y = empirical_sd,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  geom_point(
    data = results_tbl_estimation %>%
      filter(
        constraints == FALSE,
        inference == "least-squares",
        progression == "normal",
        drop_first_occasions == 0
      ),
    mapping = aes(x = n, y = median_se_bs),
    shape = 2
  ) +
  scale_y_continuous(trans = "log10", name = "Empirical SD and median estimated SE") +
  scale_x_continuous(trans = "log10", name = "Sample Size, n") +
  scale_linetype(name = "Progression Rate") +
  facet_grid(gamma_slowing ~ .)
```

In Figure \@ref(fig:coverage-bootstrap), the empirical coverage rates are
presented for the percentile confidence intervals. This reveals that there is
slight overcoverage for $n = 50$, but nominal coverage for larger sample sizes.
This indicates that the parametric bootstrap permits valid inference in small
samples.

```{r coverage-bootstrap, fig.cap="Graphs of the empirical coverage of the percentile bootstrap confidence intervals. The rows correspond to the true acceleration factor while the columns correspond to the duration of follow-up."}
results_tbl_inference %>%
  filter(constraints == FALSE,
         inference == "least-squares",
         progression == "normal",
         drop_first_occasions == 0) %>%
  ggplot(aes(
    x = n,
    y = coverage_bs,
    color = `Follow Up`,
    linetype = progression
  )) +
  geom_point() +
  geom_line() +
  geom_hline(
    # Add horizontal lines to indicate the 0.95 nominal coverage.
    data = tidyr::expand_grid(coverage = 0.95,
                              gamma_slowing = c(0.5, 0.75, 1)),
    mapping = aes(yintercept = coverage),
    alpha = 0.5
  ) +
  xlab("Sample size, n") +
  ylab(latex2exp::TeX("Empirical Coverage")) +
  scale_linetype(name = "Progression Rate") +
  facet_grid(rows = "gamma_slowing")
```

Finally, the empirical type 1 error rates and power for the parametric bootstrap
are presented in Figure \@ref(fig:error-rates-bootstrap). The corresponding 
hypothesis tests are based on the percentile confidence intervals.
First, the empirical type 1 error rate is close to nominal, or even slightly conservative,  for all sample sizes.
Second, the empirical power for meta TCT is considerably larger in ...

```{r error-rates-bootstrap, fig.cap="Graphs of the empirical type 1 error rate and power across a set of simulation settings. The hypothesis tests are based on the percentile confidence intervals. The rows correspond to the true acceleration factor."}
results_tbl_inference %>%
  filter(
    constraints == FALSE,
    inference == "least-squares",
    drop_first_occasions == 0,
    progression == "normal"
  ) %>%
  ggplot(
    aes(
      x = n,
      y = rejection_rate_bs,
      color = `Follow Up`,
      shape = `Follow Up`,
      linetype = progression
    )
  ) +
  geom_point() +
  geom_line() +
  geom_point(
    data = results_tbl_inference %>%
      filter(
        constraints == FALSE,
        inference == "least-squares",
        drop_first_occasions == 0,
        progression == "normal"
      ),
    mapping = aes(x = n,
                  y = rejection_rate_mmrm,
                  shape = `Follow Up`),
    alpha = 0.5,
    color = "gray"
  ) +
  geom_line(
    data = results_tbl_inference %>%
      filter(
        constraints == FALSE,
        inference == "least-squares",
        drop_first_occasions == 0,
        progression == "normal"
      ),
    mapping = aes(x = n,
                  y = rejection_rate_mmrm,
                  linetype = progression),
    alpha = 0.75,
    color = "gray"
  ) +
  geom_hline(
    # Add horizontal lines to indicate the 0.05 nominal error rate in the null
    # settings.
    data = tidyr::expand_grid(
      rejection_rate = 0.05,
      drop_first_occasions = 0,
      gamma_slowing = 1
    ),
    mapping = aes(yintercept = rejection_rate),
    alpha = 0.5
  ) +
  xlab("Sample size, n") +
  ylab(latex2exp::TeX("Empirical Power or Type 1 Error Rate")) +
  scale_linetype(name = "Progression Rate") +
  facet_grid(gamma_slowing ~ ., scales = "free")
```


# Summary

The results are summarized according to the goals of the simulation study.

*Primary goal: Evaluate finite sample properties of the method, also in relation
to asymptotic inference procedures.*

* Bias
* MSE, SD, SE,
* CIs and error rates

*Secondary goal: Identify settings where the meta TCT methods may yield
untrustworthy results.*

* Sample size
* Duration of follow up
* Discarding time points

*Tertiary Goal: Evaluate correctness of the implementation in the `TCT` 
R-package.*

* The functions implemented in `TCT` behaved as expected. We can thus trust the
implementation. 

# Appendix

## All Simulation Settings

In the following table, we summarize the results for all simulation settings.

```{r}
results_tbl_combined = full_join(
  x = results_tbl_estimation,
  y = results_tbl_inference
)
```

```{r}
DT::datatable(results_tbl_combined %>%
                rename("gamma_0" = "gamma_slowing") %>%
                ungroup() %>%
                mutate(across(.cols = where(is.numeric), 
                       .fns = ~ round(.x, digits = 3))), 
              caption = "Summary of all simulation scenarios.", 
              filter = "top")
```

# References




