---
title: "Analysis of the Lighthouse Data"
author: "Florian Stijven"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
params: 
  printcode: FALSE
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = params$printcode)
```


```{r setup, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(mmrm)
library(TCT)
library(gtsummary)
library(emmeans)

# Set directories where the plots and tables are saved.
dir_figures = here::here("results", "figures", "application")
```

# Introduction

In this report, we analyze the Lighthouse data. These are data from a randomized
trial in Alzheimer's disease where the ADAS-Cog score has been
evaluated at pre-determined measurement occasions.
Additionally, baseline covariates are available; these will be adjusted for to
increase efficiency.

We first read and explore the data. Next, we fit a mixed model for repeated 
measures (MMRM) and present the corresponding results. Finally, we use this 
fitted mixed model to estimate the treatment effect on the time scale with 
the meta time-component test (meta TCT) methodology.

# Data Exploration

The Lighthouse data are read via the SAS data file. These data are in a
partially long format because every post-randomization measurement has its own
row, but the baseline measurement is represented by a column. For applying meta
TCT, we need to fit a mixed model where the baseline measurement is also treated
as a response instead of a baseline covariate. Hence, we reformat the data
accordingly.

```{r}
# Read the sas-data set.
data_lighthouse = haven::read_sas(here::here("data", "original", "tct_adascog.sas7bdat"))
# The original data are not in the long format because there is no separate row
# for the ADASCog at baseline. We transform the data into the long format.
data_lighthouse_long = data_lighthouse %>%
  # Add baseline value to the change score to get the original ADAS-Cog score.
  mutate(CHG = CHG + BASVAL) %>%
  pivot_wider(names_from = "WEEK",
              values_from = "CHG",
              names_prefix = "Week_") %>%
  rename(Week_0 = "BASVAL") %>%
  pivot_longer(cols = starts_with("Week"),
               names_to = "Week",
               values_to = "ADASCog") %>%
  mutate(Week_int = as.integer(stringr::str_sub(Week, start = 6)))
```

The baseline covariates and ADAS-Cog scores are summarized by treatment group in 
Table \@ref(tab:patient-characteristics). By randomization, the distribution of
baseline covariates should be identical across treatment groups, apart from 
discrepancies due to sampling variability. This table, however, highlights two
potential issues:

1. At the post-randomization measurement occasions, there appears to be a
  discrepancy in the proportion of missing values between the treatment groups. 
2. There are centers with patients in only one of the treatment groups. This 
  does not cause problems in our further analyses. Still, this is not ideal. 

```{r patient-characteristics}
data_lighthouse_long %>%
  select(-Week_int) %>%
  pivot_wider(
    names_from = "Week",
    values_from = "ADASCog",
    names_prefix = "ADASCog"
  ) %>%
  select(-SUBJID) %>%
  tbl_summary(by = "TRT", missing_text = "(Missing)") %>%
  modify_caption("Patient Characteristics by Treatment Group") 
```

To investigate the discrepancies in the proportions of missing values, we test
whether the proportions of missing values are equal at each measurement
occasion. The results are shown in Table \@ref(tab:missingness) and clearly show
that there are more missing values in the active treatment group. The magnitude
of missingness is also substantial; 35% of the values are missing at 48 weeks in
the active treatment group. In what follows, we assume missing at random (MAR).
A MMRM leads to valid inferences under MAR (and a correctly specified model)
regardless of the proportion of missing values; however, the results could be
sensitive to small violations of MAR because there are many missing values.

```{r missingness}
data_lighthouse_long %>%
  select(-Week_int) %>%
  pivot_wider(names_from = "Week",
              values_from = "ADASCog",
              names_prefix = "ADASCog") %>%
  mutate(
    missing_Week_0 = is.na(ADASCogWeek_0),
    missing_Week_12 = is.na(ADASCogWeek_12),
    missing_Week_24 = is.na(ADASCogWeek_24),
    missing_Week_40 = is.na(ADASCogWeek_40),
    missing_Week_48 = is.na(ADASCogWeek_48)
  ) %>%
  select(
    missing_Week_12,
    missing_Week_24,
    missing_Week_40,
    missing_Week_48,
    TRT
  ) %>%
  tbl_summary(by = "TRT", missing_text = "(Missing)") %>%
  modify_caption("Comparison of Proportion of Missing Values.") %>%
  add_p(test = list(all_categorical() ~ "fisher.test")) %>%
  as_gt()
```



In Figure \@ref(fig:spaghetti-plot), the patient-specific longitudinal profiles
are plotted alongside the treatment specific mean profiles. Note that the mean
profiles are based on a complete case analysis. This figure reveals (i) that
there is a lot of between-patient variability and (ii) that there is a treatment
effect that increases with time.


```{r spaghetti-plot, fig.cap="Plot of patient-specific profiles together with the treatment specific mean profiles based on a complete case analysis.", message=FALSE}
data_lighthouse_long %>%
  na.omit() %>%
  ggplot(aes(x = Week_int, y = ADASCog, color = TRT)) +
  geom_line(aes(group = SUBJID), alpha = 0.25) +
  geom_point(
    data = data_lighthouse_long %>%
      group_by(TRT, Week, Week_int) %>%
      summarize(mean_score = mean(ADASCog, na.rm = TRUE)) %>%
      ungroup(),
    mapping = aes(x = Week_int, y = mean_score),
    color = "black"
  ) +
  geom_line(
    data = data_lighthouse_long %>%
      group_by(TRT, Week, Week_int) %>%
      summarize(mean_score = mean(ADASCog, na.rm = TRUE)) %>%
      ungroup(),
    mapping = aes(x = Week_int, y = mean_score, group = TRT),
    linewidth = 1
  ) +
  xlab("Weeks") +
  ylab("ADASCog score") +
  scale_color_discrete(name = "Treatment") +
  theme_bw()
```

# Mixed Model for Repeated Measures

We now fit a MMRM where the vector of ADAS-Cog scores at baseline, 12, 24, 40,
and 48 weeks is the response variable. The model is fitted with restricted 
maximum likelihood. We also add the following baseline covariates in the model:

* `SITEGR1`: This is a categorical variable for the study sites; we include a main
effect for this variable.
* `AAPOEN`: This is the APOE Allele status. We include an interaction between the measurement
occasion and this variable.

In this model, we also assume that the mean ADAS-Cog scores at baseline are 
equal between both treatment groups. This holds by randomization. The mixed model
is fitted with `mmrm()` from the `mmrm` R package.

```{r}
# Fit model for the full ADAS-Cog vector, including the baseline response. For
# estimating the acceleration factor, we need to assumed that the mean response
# at baseline is equal in both treatment groups. This can be "forced" by setting
# the treatment variable to "placebo" for all measurements at baseline.
data_lighthouse_long = data_lighthouse_long %>%
  mutate(
    Week = factor(Week),
    SITEGR1 = factor(SITEGR1),
    AAPOEN = factor(AAPOEN),
    TRT = ifelse(Week_int == 0L, "Placebo", TRT),
    TRT = factor(TRT)
  )
mmrm_fit_tct = mmrm(
  formula = ADASCog ~ Week:TRT + SITEGR1 + AAPOEN*Week +
    us(Week | SUBJID),
  data = data_lighthouse_long,
  method = "Kenward-Roger"
)
# summary(mmrm_fit_tct)
```

We test for a treatment effect using the F-test with the Kenward-Rogers 
degrees of freedom approximation. 

```{r, message=FALSE}
anova_mmrm = car::Anova(mmrm_fit_tct)
anova_mmrm
row_number_trt = which(rownames(anova_mmrm) ==  "Week:TRT")
p_trt_effect = anova_mmrm[["Pr(>=F)"]][row_number_trt]
```

The p-value for the F-test is `r round(p_trt_effect, 3)`. There is thus is a
significant treatment effect, but it is unclear how relevant this effect really
is. A first step towards interpreting the magnitude of the treatment effect is
to look at the estimated mean differences in ADAS-Cog scores.


Because we have included baseline covariates, we have to carefully define the
treatment group-specific mean ADAS-Cog scores at the various measurement
occasions. In what follows, we consider the estimated marginal means (EMMs, also
known as least-squares means) where the factor levels of the center variable and
the APOE variable receive weights proportional to their observed frequencies in
the data. Alternatively, we could use equal weights for these factors, meaning
that each center and APOE level receives the same weight. Which weights are most
appropriate depends on whether these factors were controlled by the researchers
and on the population one wishes to make inferences on. 

The EMMs with proportional weights and the corresponding contrasts by
measurement occasion are estimated with `emmeans()` and `contrast()` from the
`emmeans` R package; the corresponding output is printed below. These contrasts
indicate that the treatment effect increases with time. Still, it remains
difficult to appreciate the relevance of the observed treatment effect, i.e., is
a 2 or 3 points mean difference on the ADAS-Cog scale meaningful?

```{r}
# The least squares means are required for meta TCT.
ls_means_prop = emmeans::emmeans(mmrm_fit_tct, ~ TRT | Week, weights = "proportional")
ls_means_prop
```
```{r}
confint(emmeans::contrast(ls_means_prop, "pairwise"))
```


# Estimation of Acceleration Factors

In this section, we will use the EMMs from the previous section to estimate the
treatment effect on the time scale (i.e., the time-specific acceleration factors
and the common acceleration factor). We use the estimators implemented in the
`TCT` R package for this.

In the first step, we need to extract the EMMs at all
treatment-measurement occasion combinations. We also need the corresponding
covariance matrix. These quantities can be obtained from the `emmGrid` object that
was returned by `emmeans()`.

```{r}
# We first need to extract the EMMs for all treatment time combinations and
# the corresponding covariance matrix.
ctrl_indices = which(summary(ls_means_prop)$TRT == "Placebo")
exp_indices = which(summary(ls_means_prop)$TRT == "Active")
ctrl_estimates = summary(ls_means_prop)$emmean[ctrl_indices]
exp_estimates = summary(ls_means_prop)$emmean[exp_indices]
# Covariance matrix' rows and columns should be ordered by (i) treatment group
# and (ii) time.
vcov_emmeans = vcov(ls_means_prop)[c(ctrl_indices, exp_indices), c(ctrl_indices, exp_indices)]
```

In the second step, we use `TCT_meta()` to estimate the time-specific acceleration factors; the estimated acceleration factors are
shown below. Note that the z-values for the estimated acceleration factors are
equal to the corresponding t-values in the EMM contrasts. This is no coincidence
and always holds for the `"contrast"` tests for the time-specific acceleration
factors. The p-values below and the p-values for the EMM contrasts differ
slightly because `TCT_meta()` uses the standard normal distribution instead of
the t-distribution to compute the p-value.

```{r}
TCT_fit = TCT_meta(
  time_points = c(0, 12, 24, 40, 48),
  ctrl_estimates = ctrl_estimates,
  exp_estimates = exp_estimates,
  vcov = vcov_emmeans,
  interpolation = "spline",
  inference = "contrast",
  B = 0
)
summary(TCT_fit)
```

The estimated measurement occasion-specific acceleration factors do not seem to
contradict the assumption of proportional slowing, i.e., the assumption that the
acceleration factor is constant over time.Indeed, the estimates are all within
one standard error of 0.50. Therefore, we may proceed to estimating the common
acceleration factor. (Later on, we formally test this assumption.)

The object returned by `TCT_meta()` can be used as argument to
`TCT_meta_common()` to estimate the common acceleration factor. We use two
estimation methods: (i) the contrast-based estimator with adaptively selected weights and
(ii) the least-squares estimator.

```{r, cache=TRUE}
# Estimate common accelerations factors.
set.seed(1)
TCT_common_fit_contrast = TCT_meta_common(TCT_fit, B = 2e3, inference = "contrast")
summary(TCT_common_fit_contrast)
set.seed(1)
TCT_common_fit_ls = TCT_meta_common(TCT_fit, B = 2e3, inference = "least-squares")
summary(TCT_common_fit_ls)
```

These two estimators lead to very similar results. The estimated common
acceleration factor is about 0.50. The estimated treatment effect corresponds to
slowing disease progression with 50\% -- a very relevant effect. The p-values
and confidence intervals vary a bit more between these estimators and also
depend on whether the parametric bootstrap is used. Since the sample size is
relatively small, the parametric bootstrap should be preferred. Indeed,
simulations revealed that inference based on the parametric bootstrap is closer
to nominal in small sample sizes.

The p-value of the goodness-of-fit test (based on the minimized least-squares
criterion) for the proportional slowing assumption is 
`r round(summary(TCT_common_fit_ls)$proportional_slowing_test[["p_value"]], 3)`.
This indicates no evidence against the assumption of proportional slowing;
although, this assumption as expected to hold only approximately at best. A
larger sample size would likely lead to significant p-values.

```{r}
# Plot the estimated marginal means together with the 95% confidence intervals and
# add the natural cubic spline interpolation to get the estimated reference
# trajectory.

# Calling summary on the emmGrid object returns a data frame with the
# estimated marginal means, the corresponding standard errors and confidence limits.
emmeans_tbl = summary(ls_means_prop)
# We need to convert the (character) Week variable to an integer.
emmeans_tbl = emmeans_tbl %>%
  mutate(Week_int = as.integer(stringr::str_sub(Week, start = 6)))

# Extract the estimated marginal means for the control and active-treatment
# groups.
ref_means = emmeans_tbl %>%
  filter(TRT == "Placebo") %>%
  arrange(Week_int) %>%
  pull(emmean)
trt_means = emmeans_tbl %>%
  filter(TRT == "Active") %>%
  arrange(Week_int) %>%
  pull(emmean)
# Extract the times corresponding to the fixed measurement occasions and sort in
# increasing order.
time_points = unique(emmeans_tbl$Week_int) %>% sort()
# Define the grid on which we evaluate the spline interpolation.
time_grid = seq(from = 0,
                to = max(time_points),
                length.out = 5e3)
# Interpolate the reference trajectory using a natural cubic spline. This
# corresponds to the estimated reference trajectory for the contrast-based
# estimator of the common acceleration factor (since the control group
# parameters are no re-estimated here).
trajectory_points = spline(
  x = time_points,
  y = ref_means,
  xout = time_grid,
  method = "natural"
)$y
# Extract the re-estimates of the control group parameters.
alpha_est = TCT_common_fit_ls$alpha_est
# Interpolate between the re-estimated control group parameters. This is the
# estimated reference trajectory for the least-squares estimator of the common
# acceleration factor.
trajectory_points_ls = spline(
  x = time_points,
  y = alpha_est,
  xout = time_grid,
  method = "natural"
)$y

# Adaptive weights-based estimate of the common acceleration factor.
gamma_est_contrast = coef(TCT_common_fit_contrast)
# Least-squares estimate of the common acceleration factor.
gamma_est_ls = coef(TCT_common_fit_ls)
# Estimated trajectory for the active treatment group based on the adaptive
# weights-based estimate of the common acceleration factor.
trt_trajectory_points_contrast = spline(
  x = time_points,
  y = ref_means,
  xout = time_grid * gamma_est_contrast,
  method = "natural"
)$y
# Estimated trajectory for the active treatment group based on the least-squares
# estimate of the common acceleration factor.
trt_trajectory_points_ls = spline(
  x = time_points,
  y = alpha_est,
  xout = time_grid * gamma_est_ls,
  method = "natural"
)$y

plot_contrast = tibble(time = time_grid, y = trajectory_points, TRT = "Placebo") %>%
  bind_rows(tibble(time = time_grid, y = trt_trajectory_points_contrast, TRT = "Active")) %>%
  ggplot(aes(x = time, y = y, color = TRT)) +
  geom_line(alpha = 0.5) +
  geom_point(
    data = emmeans_tbl,
    aes(x = Week_int, y = emmean, color = TRT)
  ) +
  # Add confidence interval limits.
  geom_errorbar(
    data = emmeans_tbl,
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.5,
    width = 1
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

plot_ls = tibble(time = time_grid, y = trajectory_points_ls, TRT = "Placebo") %>%
  bind_rows(tibble(time = time_grid, y = trt_trajectory_points_ls, TRT = "Active")) %>%
  ggplot(aes(x = time, y = y, color = TRT)) +
  geom_line(alpha = 0.5) +
  geom_point(
    data = emmeans_tbl,
    aes(x = Week_int, y = emmean, color = TRT)
  ) +
  # Add confidence interval limits.
  geom_errorbar(
    data = emmeans_tbl,
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.5,
    width = 1
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

# Paste the plots together. There should be one common legend.
library(patchwork)
plot_contrast + plot_ls +
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(tag_levels = "A")
  

ggsave(
  filename = "estimated-trajectory.pdf",
  path = dir_figures,
  device = "pdf",
  width = double_width,
  height = single_height,
  units = "cm",
  dpi = res
)

# Also save the plots separately.
ggsave(
  plot = plot_contrast,
  filename = "estimated-trajectory-contrast.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

ggsave(
  plot = plot_ls,
  filename = "estimated-trajectory-ls.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)
```

```{r}
# Additional plots, useful for presentations etc.

# Plot of the estimated time-sepcific means with 95% confidence intervals. No
# interpolation is added to emphasize the default analysis method of
# time-specific mean differences.
plot_means = emmeans_tbl %>% 
  ggplot(aes(x = time, y = y, color = TRT)) +
  geom_point(
    aes(x = Week_int, y = emmean, color = TRT)
  ) +
  # Add confidence interval limits.
  geom_errorbar(
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.5,
    width = 1
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

ggsave(
  plot = plot_means,
  filename = "estimated-means.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

ggsave(
  plot = plot_means,
  filename = "estimated-means.png",
  path = dir_figures,
  device = "png",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

# Plot the reference trajectory without any time mappings.
plot_reference_trajectory = tibble(time = time_grid, y = trajectory_points, TRT = "Placebo") %>%
  ggplot(aes(x = time, y = y)) +
  geom_line(aes(color = TRT), alpha = 0.5) +
  geom_point(data = emmeans_tbl,
             aes(x = Week_int, y = emmean, color = TRT),
             alpha = 0.50) +
  # Add confidence interval limits.
  geom_errorbar(
    data = emmeans_tbl,
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.25,
    width = 1
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

ggsave(
  plot = plot_reference_trajectory,
  filename = "reference-trajectory.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

ggsave(
  plot = plot_reference_trajectory,
  filename = "reference-trajectory.png",
  path = dir_figures,
  device = "png",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

# Plot both the reference trajectory and the estimated trajectory for the active
# treatment group based on the contrast-based estimator of the common
# acceleration factor.
plot_contrast_trajectories = tibble(time = time_grid, y = trajectory_points, TRT = "Placebo") %>%
  bind_rows(tibble(time = time_grid, y = trt_trajectory_points_contrast, TRT = "Active")) %>%
  ggplot(aes(x = time, y = y, color = TRT)) +
  geom_line(alpha = 0.5) +
  geom_point(
    data = emmeans_tbl,
    aes(x = Week_int, y = emmean, color = TRT)
  ) +
  # Add confidence interval limits.
  geom_errorbar(
    data = emmeans_tbl,
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.5,
    width = 1
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

ggsave(
  plot = plot_contrast_trajectories,
  filename = "contrast-trajectories.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

ggsave(
  plot = plot_contrast_trajectories,
  filename = "contrast-trajectories.png",
  path = dir_figures,
  device = "png",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

# Plot the reference trajectory and the time mappings with arrows from the fixed
# measurement occasions to the mapped times.

# Compute time mappings.
time_mappings = tibble(
  time = c(12, 24, 40, 48),
  mapped_time = c(12, 24, 40, 48) * coef(TCT_fit),
  outcome_value = spline(
    x = time_points,
    y = ref_means,
    xout = mapped_time,
    method = "natural"
  )$y,
  lower_ci_mapped_time = c(12, 24, 40, 48) * summary(TCT_fit)$ci_matrix[, 1],
  upper_ci_mapped_time = c(12, 24, 40, 48) * summary(TCT_fit)$ci_matrix[, 2]
)

plot_contrast_time_mappings_arrows = tibble(time = time_grid, y = trajectory_points, TRT = "Placebo") %>%
  ggplot(aes(x = time, y = y)) +
  geom_line(aes(color = TRT), alpha = 0.5) +
  geom_point(data = emmeans_tbl,
             aes(x = Week_int, y = emmean, color = TRT),
             alpha = 0.50) +
  # Add confidence interval limits.
  geom_errorbar(
    data = emmeans_tbl,
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.25,
    width = 1
  ) +
  geom_segment(
    data = time_mappings,
    aes(
      x = time,
      xend = mapped_time,
      y = outcome_value,
      yend = outcome_value
    ),
    arrow = arrow(length = unit(0.25, 'cm'))
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

ggsave(
  plot = plot_contrast_time_mappings_arrows,
  filename = "time-mappings-arrows.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

ggsave(
  plot = plot_contrast_time_mappings_arrows,
  filename = "time-mappings-arrows.png",
  path = dir_figures,
  device = "png",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

# Plot the reference trajectory and the time mappings with 95% CIs around the
# mapped times.
plot_contrast_time_mappings = tibble(time = time_grid, y = trajectory_points, TRT = "Placebo") %>%
  ggplot(aes(x = time, y = y)) +
  geom_line(aes(color = TRT), alpha = 0.5) +
  geom_point(data = emmeans_tbl,
             aes(x = Week_int, y = emmean, color = TRT),
             alpha = 0.50) +
  # Add confidence interval limits.
  geom_errorbar(
    data = emmeans_tbl,
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.25,
    width = 1
  ) +
  # Add time mappings with confidence intervals.
  geom_point(data = time_mappings, aes(x = mapped_time, y = outcome_value)) +
  geom_errorbarh(
    data = time_mappings,
    aes(y = outcome_value, xmin = lower_ci_mapped_time, xmax = upper_ci_mapped_time),
    height = 0.25
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

ggsave(
  plot = plot_contrast_time_mappings,
  filename = "time-mappings.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

ggsave(
  plot = plot_contrast_time_mappings,
  filename = "time-mappings.png",
  path = dir_figures,
  device = "png",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)
```

```{r}
# Simulate five patient-level trajectories by pertubring the estimated reference
# trajectory with a random vertical shift and random acceleration factor.
set.seed(1)
simulated_trajectories = tibble(time = time_grid, y = trajectory_points) %>%
  slice(rep(1:n(), each = 5)) %>%
  mutate(patient_id = rep(1:5, times = n() / 5)) %>%
  group_by(patient_id) %>%
  mutate(
    vertical_shift = rnorm(1, mean = 0, sd = 0.3),
    acceleration_factor = rnorm(1, mean = 1, sd = 0.2)
  ) %>%
  group_by(time) %>%
  mutate(
    y_simulated = spline(
      x = time_points,
      y = ref_means,
      xout = time * acceleration_factor,
      method = "natural"
    )$y + vertical_shift
  ) %>%
  ungroup() %>%
  mutate(
    TRT = "Placebo"
  )
# Add active treatment group trajectories by applying the estimated common
# acceleration factor to the time variable.
simulated_trajectories = simulated_trajectories %>%
  bind_rows(
    tibble(time = time_grid, y = trajectory_points) %>%
      slice(rep(1:n(), each = 5)) %>%
      mutate(patient_id = (rep(1:5, times = n(
      ) / 5) + 5)) %>%
      group_by(patient_id) %>%
      mutate(
        vertical_shift = rnorm(1, mean = 0, sd = 0.2),
        acceleration_factor = rnorm(1, mean = 1, sd = 0.2)
      ) %>%
      group_by(time) %>%
      mutate(
        y_simulated = spline(
          x = time_points,
          y = ref_means,
          xout = time * acceleration_factor * coef(TCT_common_fit_contrast),
          method = "natural"
        )$y + vertical_shift
      ) %>%
      ungroup() %>%
      mutate(TRT = "Active")
  )

# Plot the simulated trajectories together with the estimated reference
# trajectory.
plot_simulated_trajectories = ggplot() +
  geom_line(data = simulated_trajectories,
            aes(x = time, y = y_simulated, group = patient_id, color = TRT),
            alpha = 0.5) +
  geom_point(data = emmeans_tbl,
             aes(x = Week_int, y = emmean, color = TRT),
             alpha = 0.50) +
  # Add confidence interval limits.
  geom_errorbar(
    data = emmeans_tbl,
    aes(
      x = Week_int,
      ymin = lower.CL,
      ymax = upper.CL,
      y = emmean,
      color = TRT
    ),
    alpha = 0.25,
    width = 1
  ) +
  scale_x_continuous(name = "Weeks Since Randomization", breaks = c(time_points)) +
  scale_y_continuous(name = expr(hat(E)(Y[t](z)))) +
  scale_color_brewer(type = "qual",
                     palette = 2,
                     name = "Treatment")

ggsave(
  plot = plot_simulated_trajectories,
  filename = "simulated-trajectories.pdf",
  path = dir_figures,
  device = "pdf",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)

ggsave(
  plot = plot_simulated_trajectories,
  filename = "simulated-trajectories.png",
  path = dir_figures,
  device = "png",
  width = single_width,
  height = single_height,
  units = "cm",
  dpi = res
)
```



# Conclusion

Using the classic analysis method, the MMRM and the F-test, we found a
significant treatment effect, p = `r round(p_trt_effect, 3)`. Based on contrasts
in EMMs, the estimated treatment effects on the ADAS-Cog scale at 40 and 48
weeks are a mean difference of 
`r round(summary(contrast(ls_means_prop, "pairwise"))$estimate[4], 2)` and 
`r round(summary(contrast(ls_means_prop, "pairwise"))$estimate[5], 2)` points, 
respectively. The relevance of these treatment effects is unclear. The estimated
common acceleration factor is
`r round(coef(TCT_common_fit_ls), 2)`, p = `r round(summary(TCT_common_fit_ls)$p_bootstrap, 3)`
(using the least-squares estimator and parametric bootstrap). The meta TCT
approach thus reveals a clinically very relevant effect, something the MMRM did
not reveal. In addition, the p-value is considerably smaller as compared to
F-test in the MMRM.

