---
title: "Analysis of the Lighthouse Data"
author: "Florian Stijven"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
params: 
  printcode: FALSE
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = params$printcode)
```


```{r setup, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(mmrm)
library(TCT)
library(gtsummary)
library(emmeans)
```

# Introduction

In this report, we analyze the mock data that resemble the data in the case
study. These are data from a randomized trial in Alzheimer's disease where the
ADAS-Cog score has been evaluated at pre-determined measurement occasions.
Additionally, baseline covariates are available; these will be adjusted for to
increase efficiency.

We first read and explore the data. Next, we fit a mixed model for repeated 
measures (MMRM) and present the corresponding results. Finally, we use this 
fitted mixed model to estimate the treatment effect on the time scale with 
the meta time-component test (meta TCT) methodology.

# Data Exploration

The Lighthouse data are read via the SAS data file. These data are in a
partially long format because every post-randomization measurement has its own
row, but the baseline measurement is represented by a column. For applying meta
TCT, we need to fit a mixed model where the baseline measurement is also treated
as a response instead of a baseline covariate. Hence, we reformat the data
accordingly.

```{r}
# Read the sas-data set and convert the variables to a format that closely
# resembles the real data in the case study.
data_lighthouse = haven::read_sas("application/mock-analysis/effect.sas7bdat") %>%
  rename(WEEK = time, SUBJID = id, TRT = trt) %>%
  mutate(
    WEEK = case_match(WEEK, 1 ~ 12, 2 ~ 24, 3 ~ 40, 4 ~ 48),
    TRT = case_match(TRT, 0 ~ "Placebo", 1 ~ "Active"),
    SITE = factor(SITE)
  )
# The data are not in the long format because there is no separate row for the
# ADASCog score at baseline. We recode the data into the long format.
data_lighthouse_long = data_lighthouse %>%
  # The sas-data set contains the change scores as outcome variable, instead of
  # the ADASCog score itself. Add baseline value to the change score to get the
  # original ADASCog score.
  mutate(CHG = CHG + BASVAL) %>%
  pivot_wider(names_from = "WEEK",
              values_from = "CHG",
              names_prefix = "Week_") %>%
  rename(Week_0 = "BASVAL") %>%
  pivot_longer(cols = starts_with("Week"),
               names_to = "Week",
               values_to = "ADASCog") %>%
  mutate(Week_int = as.integer(stringr::str_sub(Week, start = 6)))
```

The baseline covariates and ADAS-Cog scores are summarized by treatment group in 
Table \@ref(tab:patient-characteristics). By randomization, the distribution of
baseline covariates should be identical across treatment groups, apart from 
discrepancies due to sampling variability. 

```{r patient-characteristics}
data_lighthouse_long %>%
  select(-Week_int) %>%
  pivot_wider(
    names_from = "Week",
    values_from = "ADASCog",
    names_prefix = "ADASCog"
  ) %>%
  select(-SUBJID) %>%
  tbl_summary(by = "TRT", missing_text = "(Missing)") %>%
  modify_caption("Patient Characteristics by Treatment Group") 
```

In Figure \@ref(fig:spaghetti-plot), the patient-specific longitudinal profiles
are plotted alongside the treatment specific mean profiles. Note that the mean
profiles are based on a complete case analysis. This figure reveals (i) that
there is a lot of between-patient variability and (ii) that there is a treatment
effect that increases with time, although the effect seems negligible.


```{r spaghetti-plot, fig.cap="Plot of patient-specific profiles together with the treatment specific mean profiles based on a complete case analysis.", message=FALSE}
data_lighthouse_long %>%
  na.omit() %>%
  ggplot(aes(x = Week_int, y = ADASCog, color = TRT)) +
  geom_line(aes(group = SUBJID), alpha = 0.25) +
  geom_point(
    data = data_lighthouse_long %>%
      group_by(TRT, Week, Week_int) %>%
      summarize(mean_score = mean(ADASCog, na.rm = TRUE)) %>%
      ungroup(),
    mapping = aes(x = Week_int, y = mean_score),
    color = "black"
  ) +
  geom_line(
    data = data_lighthouse_long %>%
      group_by(TRT, Week, Week_int) %>%
      summarize(mean_score = mean(ADASCog, na.rm = TRUE)) %>%
      ungroup(),
    mapping = aes(x = Week_int, y = mean_score, group = TRT),
    linewidth = 1
  ) +
  xlab("Weeks") +
  ylab("ADASCog score") +
  scale_color_discrete(name = "Treatment") +
  theme_bw()
```

# Mixed Model for Repeated Measures

We now fit a MMRM where the vector of ADAS-Cog scores at baseline, 12, 24, 40,
and 48 weeks is the response variable. The model is fitted with restricted 
maximum likelihood. We also add the following baseline covariates in the model:

* `SITE`: This is a categorical variable for the study sites; we include a main
effect for this variable.
* `AAPOEN`: This is the APOE Allele status. We include an interaction between the measurement
occasion and this variable.

In this model, we also assume that the mean ADAS-Cog scores at baseline are 
equal between both treatment groups. This holds by randomization. The mixed model
is fitted with `mmrm()` from the `mmrm` R package.

```{r}
# Fit model for the full ADASCog vector, including the baseline response. For
# the TCT methods, it is most ideal to assume that the mean response at baseline
# is equal in both treatment groups. This can be "forced" by setting the
# treatment variable to "placebo" for all measurements at baseline.
data_lighthouse_long = data_lighthouse_long %>%
  mutate(
    Week = factor(Week),
    AAPOEN = factor(AAPOEN),
    TRT = ifelse(Week_int == 0L, "Placebo", TRT),
    TRT = factor(TRT),
    SUBJID = factor(SUBJID)
  )
mmrm_fit_tct = mmrm(
  formula = ADASCog ~ Week:TRT + SITE + AAPOEN*Week +
    us(Week | SUBJID),
  data = data_lighthouse_long,
  method = "Kenward-Roger"
)
# summary(mmrm_fit_tct)
```

We test for a treatment effect using the F-test with the Kenward-Rogers 
degrees of freedom approximation. 

```{r, message=FALSE}
anova_mmrm = car::Anova(mmrm_fit_tct)
anova_mmrm
row_number_trt = which(rownames(anova_mmrm) ==  "Week:TRT")
p_trt_effect = anova_mmrm[["Pr(>=F)"]][row_number_trt]
```

The p-value for the F-test is `r round(p_trt_effect, 3)`. There is thus is a 
significant treatment effect, but it is unclear how relevant this effect really 
is. A first step towards interpreting the magnitude of the treatment effect is to look
at the estimated mean differences in ADAS-Cog scores.


Because we have included baseline covariates, we have to carefully define the
treatment group-specific mean ADAS-Cog scores at the various measurement
occasions. In what follows, we consider the estimated marginal means (EMMs, also
known as least-squares means) where the factor levels of the center variable and
the APOE variable receive weights proportional to their observed frequencies in
the data. Alternatively, we could use equal weights for these factors, meaning
that each center and APOE level receives the same weight. Which weights are most
appropriate depends on whether these factors were controlled by the researchers
and on the population one wishes to make inferences on. 

The EMMs with proportional weights and the corresponding contrasts by
measurement occasion are estimated with `emmeans()` and `contrast()` from the
`emmeans` R package; the corresponding output is printed below. These contrasts
indicate that the treatment effect increases with time. Still, it remains
difficult to appreciate the relevance of the observed treatment effect, i.e., is
a 2 points mean difference on the ADAS-Cog scale meaningful?

```{r}
# The least squares means are required for meta TCT.
ls_means_prop = emmeans::emmeans(mmrm_fit_tct, ~ TRT | Week, weights = "proportional")
ls_means_prop
```
```{r}
emmeans::contrast(ls_means_prop, "pairwise")
```


# Meta Time-Component Test

In this section, we will use the EMMs from the previous section to estimate the
treatment effect on the time scale. We use the meta TCT methodology as implemented
in the `TCT` R package for this. 

In the first step, we need to extract the EMMs at all
treatment-measurement occasion combinations. We also need the corresponding
covariance matrix. These quantities can be obtained from the `emmGrid` object that
was returned by `emmeans()`.

```{r}
# We first need to extract the LS means for all treatment time combinations and
# the corresponding covariance matrix.
ctrl_indices = which(summary(ls_means_prop)$TRT == "Placebo")
exp_indices = which(summary(ls_means_prop)$TRT == "Active")
ctrl_estimates = summary(ls_means_prop)$emmean[ctrl_indices]
exp_estimates = summary(ls_means_prop)$emmean[exp_indices]
# Covariance matrix' rows and columns should be ordered by (i) treatment group
# and (ii) time.
vcov_emmeans = vcov(ls_means_prop)[c(ctrl_indices, exp_indices), c(ctrl_indices, exp_indices)]
```

In the second step, we use `TCT_meta()` to estimate the measurement
occasion-specific acceleration factors; the estimated acceleration factors are
shown below. Note that the z-values for the estimated acceleration factors are
equal to the corresponding t-values in the EMM contrasts. This is no coincidence
and always holds for the `"score"` tests for the time-specific acceleration
factors. The p-values below and the p-values for the EMM contrasts differ
slightly because `TCT_meta()` uses the standard normal distribution instead of
the t-distribution to compute the p-value.

```{r}
TCT_fit = TCT_meta(
  time_points = c(0, 12, 24, 40, 48),
  ctrl_estimates = ctrl_estimates,
  exp_estimates = exp_estimates,
  vcov = vcov_emmeans,
  interpolation = "spline",
  inference = "score",
  B = 0
)
summary(TCT_fit)
```

The estimated measurement occasion-specific acceleration factors do not
contradict the assumption of proportional slowing, i.e., the assumption that the
acceleration factor is constant over time. Indeed, the estimates are all within
one standard error of 0.50. Therefore, we can proceed to estimating the common
acceleration factor.

The object returned by `TCT_meta()` can be used as argument to
`TCT_meta_common()` to estimate the common acceleration factor. We use two
estimation methods: (i) the score estimator with adaptively selected weights and
(ii) the least-squares estimator.

```{r, cache=TRUE}
# Estimate common accelerations factors.
set.seed(1)
TCT_common_fit_score = TCT_meta_common(TCT_fit,
                                       B = 1e3,
                                       inference = "score",
                                       type = "custom")
summary(TCT_common_fit_score)
set.seed(1)
TCT_common_fit_ls = TCT_meta_common(TCT_fit,
                                    B = 1e3,
                                    inference = "least-squares")
summary(TCT_common_fit_ls)
```

These two estimators lead to very similar results. The estimated common
acceleration factor is about 0.77 or 0.80. The estimated treatment effect
corresponds to slowing disease progression with about 20\% -- a very relevant
effect. The p-values and confidence intervals vary a bit more between these
estimators and also depend on whether the parametric bootstrap is used. Since
the sample size is relatively small, the parametric bootstrap should be
preferred. Indeed, simulations revealed that inference based on the parametric
bootstrap is closer to nominal in small sample sizes.

# Conclusion

Using the classic analysis method, the MMRM and the F-test, we found a
significant treatment effect, p = `r round(p_trt_effect, 3)`. Based on contrasts
in EMMs, the estimated treatment effects on the ADAS-Cog scale at 40 and 48
weeks are a mean difference of 
`r round(summary(contrast(ls_means_prop, "pairwise"))$estimate[4], 2)` and 
`r round(summary(contrast(ls_means_prop, "pairwise"))$estimate[5], 2)` points, 
respectively. The relevance of these treatment effects is unclear. The estimated
common acceleration factor is
`r round(coef(TCT_common_fit_ls), 2)`, p = `r round(summary(TCT_common_fit_ls)$p_bootstrap, 3)`
(using the least-squares estimator and parametric bootstrap). The meta TCT
approach thus reveals a clinically modest effect.The corresponding 95 \% CI is 
`r paste0("(", round(summary(TCT_common_fit_ls)$ci_bootstrap[1], 2), ", ", round(summary(TCT_common_fit_ls)$ci_bootstrap[2], 2), ")")`, 
which reveals that clinically large treatment effects -- e.g., slowing of more
than 50 \% -- can be excluded. In addition, the p-value is modestly smaller as
compared to F-test in the MMRM.


